Flowchart of Current Algorithm

### 2009-10-19 docs/flowchart/flowchart.tex
chopped out outdated material that appears in the multilevel paper.  this can be added later

Here we outline the algorithm currently implemented in the code.  Our
starting point for this description are the series of papers describing
the development of the algorithm:
\begin{itemize}
\item {\em Low Mach Number Modeling of Type Ia
  Supernovae. I. Hydrodynamics,} A.~S.~Almgren, J.~B.~Bell, 
  C.~A.~Rendleman, \& M.~Zingale 2006, ApJ, 637, 922 (henceforth
  paper~I)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
  Evolution,} A.~S.~Almgren, J.~B.~Bell, C.~A.~Rendleman, \& M.~Zingale
  2006, ApJ, 649, 927 (henceforth paper~II)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
  A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& M.~Zingale
  2008, ApJ, 684, 449 (henceforth paper~III)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
  M.~Zingale, A.~S.~Almgren, J.~B.~Bell, A.~Nonaka, \& S.~E.~Woosley
  2009, ApJ, 704, 196 (henceforth paper~IV)
\item {\em MAESTRO: An Adaptive Low Mach Number Hydrodynamics Algorithm for Stellar
  Flows,} A.~Nonaka, A.~S.~Almgren, J.~B.~Bell, M.~J.~Lijewski, C.~M.~Malone,
  \& M.~Zingale 2009, {\it Submited to SIAM J. Sci. Comput.}
  (henceforth ``the multilevel paper'')
\end{itemize}
We also carry over some ideas from our small-scale low Mach number algorithm
for astrophysical flows:
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J.~B.~Bell, M.~S.~Day, C.~A.~Rendleman, S.~E.~Woosley, \& M.~Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

\section{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item {\tt fill\_3d} and {\tt average} algorithms.
\item We update temperature after each {\bf React State} as well as the advection step.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
than $T=T(\rho,h,X_k)$.
Changes which modify enthalpy update.  These are separate because in this paper,
enthalpy is only used as a diagnostic since we use {\tt use\_tfromp}.  If we don't
mention these items in paper IV, they should me noted in paper 4.5.
\begin{enumerate}
\item We have changed the discretization of $\Ubt\cdot\nabla p_0$ in the enthalpy
update to $\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
\item In paper III we discretized the enthalpy evolution equation in
terms of $T$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\section{Changes Between Papers 4 and Papers 4.5}
\begin{enumerate}
\item We split the update of the density and enthalpy in steps 4 and 8.
Now, density is updated first, allowing us to compute time-centered values of 
$\psi$ that are then used in the enthalpy update.
\item We split {\bf Advect Base} into separate {\bf Advect Base Density},
{\bf Advect Base Pressure}, and {\bf Advect Base Enthalpy}.  The calling sequence
is in the main flow chart.
\item We have eliminated {\bf Advect Base Pressure} entirely, and instead directly
enforce HSE after {\bf Correct Base}.  {\bf Advect Base Pressure} was just a fancy
way of enforcing HSE anyway, so we just compute the new pressure directly.  The
problems we have seen with enforcing HSE directly may have to do with the roundoff
error that accumulates, causing the temperature computed from the EOS to drift.
However, using {\bf Advect Base Pressure}, the latest version of {\bf test\_basestate}
generates negative pressures, so that is even worse.
\item We replaced {\bf Correct Base} with $\rho_0 = {\rm{\bf Avg}}(\rho)$ in 
planar geometry and $\rho_0 = \rho_0 + {\rm{\bf Avg}}(\rho - \rho_0)$ in spherical geometry,
eliminating the need for $\nabla\cdot\etarho$.  {\bf Correct Base} is now used
analogously for setting $(\rho h)_0$ after {\bf React State}.
\item After {\bf React State}, we set $(\rho h)_0 = {\rm{\bf Avg}}(\rho h)$ in 
planar geometry and $(\rho h)_0 = (\rho h)_0 + {\rm{\bf Avg}}[(\rho h) - (\rho h)_0]$ 
in spherical geometry.  This means that {\bf React Base} no longer exists, and the 
intermediate states, $(\rho h)_0^{(1)}$ and $(\rho h)_0^{(2)}$, no longer have to exist.  
{\bf Advect Base Enthalpy} will advance $(\rho h)_0^n \rightarrow (\rho h)_0^{n+1}$ 
directly.  This also means we don't have to deal with $\eta_{\rho h}$ or a 
{\bf Diffuse Base}.
\item In {\bf Step 4} and {\bf Step 8}, we moved {\bf Correct Base} so it
occurs after the full state density update and  before {\bf Advect Base Pressure}.  
Since the full enthalpy advance ``feels'' the effects of the time-centered $\etarho$, 
we should rely on a $\rho_0^{n+1}$ that also feels the effects.
\item {\bf AMR!!!}
Changes specific to spherical geometry:
\begin{enumerate}
\item Discuss {\bf Advect Base} routines and {\bf Correct Base} routine.
\item $\etarho$ and $\psi$ equations.
\section{Changes Between Papers 4.5 and Paper 5}
\begin{enumerate}
\item Use PPM limiters in advection.
\section{Changes Between Paper 5 and XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.

@@

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \eb_r$) & cm~s$^{-1}$ \\
\hline
$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv \overline{(\rho' \Ub \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}
%
\renewcommand{\arraystretch}{1.0}
Here are some charts demonstrating what the superscript notation above scalars implies:
\begin{equation}
\left.\begin{array}{ccccccccc}
& \text{React State} & & \text{Advect State} & & \text{Diffuse Enthalpy} & & \text{React State} & \\
\rho^n & = & \rho^{(1)} & \longrightarrow & & & \rho^{(2)} & = & \rho^{n+1} \\
(\rho X_k)^n & \longrightarrow & (\rho X_k)^{(1)} & \longrightarrow & & & (\rho X_k)^{(2)} & \longrightarrow & (\rho X_k)^{n+1} \\
(\rho h)^n & \longrightarrow & (\rho h)^{(1)} & \longrightarrow & (\rho h)^{(1a)} & \longrightarrow & (\rho h)^{(2)} & \longrightarrow & (\rho h)^{n+1}\\
T^n & \longrightarrow & T^{(1)} & & & \longrightarrow & T^{(2)} & \longrightarrow & T^{n+1} \\
\end{array}\right.\nonumber
\end{equation}
\begin{equation}
\left.\begin{array}{ccccccc}
& \text{Advect Base} & & \text{Correct Base} & \\
\rho_0^n & \longrightarrow & \rho_0^{(2a)} & \longrightarrow & \rho_0^{n+1} \\
(\rho h)_0^n & \longrightarrow & & & (\rho h)_0^{n+1} \\
p_0^n & \longrightarrow & & & p_0^{n+1} \\
\end{array}\right.\nonumber
\end{equation}
\section{Shorthand Functions}
We make use of the following shorthand notations in outlining the algorithm:
\begin{enumerate}

\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
as the average over a layer at constant radius r, $\Omega_H$, as
\begin{equation}
\overline{\phi(r)}
= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}
\begin{description}
\item[planar:] This is a straightforward average since the radial coordinate is 
in alignment with the lateral coordinate.
\item[spherical:] See the description in paper IV, section 2.2 
(or section \ref{Sec:Avg} in these notes).
\end{description}

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}, 
p_0^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, 
(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite vode).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
from equation (8) in paper III,
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We update temperature using
$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
$T^{\outp} = T(\rho^\outp,p_0^\inp,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
Note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

After the new mass fractions have been computed, the reaction rates are defined as:
\begin{equation}
(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
\end{equation}
and the nuclear energy generation rate is defined as
\begin{equation}
(\rho\Hnuc)^{\outp} = -\sum_k q_k (\rho\omegadot_k)^{\outp} \enskip,
\end{equation}
for the case where only strong-mediated reactions are involved.  As
$\Hnuc$ is an output of the reaction network, more general networks
(i.e., involving weak interactions) can easily be added.  The enthalpy
update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
updated by
\begin{equation}
(\rho h)^{\outp} = 
(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
\end{equation}

\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
$\rightarrow [\rho_0^\outp,\rho_0^{\outp,n+\myhalf}]$
is the process by which we update the base state density through $\dt$ in time
as well as compute time-centered interface states, $\rho_0^{\outp,n+\myhalf}$, 
for use later in the algorithm.
\begin{description}

\item[planar:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nabla \cdot \left( \rhozero w_0 \eb_r \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ] \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

\item[spherical:] The base state density update now includes the area factors in the 
divergences:
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
          \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right] \enskip. 
\end{equation}
%\item {\bf Advect Base Pressure}
%is the process by which we update the base state pressure through $\dt$ in time and
%compute $\psi$.  The input arguments are different for planar and spherical geometries.
%\begin{description}
%\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
%[p_0^\outp,\psi^\outp]$:
%
%First we compute a cell-centered $\psi^\outp$ using
%\begin{equation}
%\psi_j^\outp = 
%\frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
%\end{equation}
%Then, we discretize the base state pressure evolution equation from paper III equation (31),
%\begin{equation}
%\frac{\partial p_0}{\partial t} = 
%-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip , \label{eq:p0upd_new}
%\end{equation}
%to compute the new base state pressure,
%\begin{equation}
%p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
%+ w_{0,j-\myhalf}^\inp \right)
%\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) 
%+ \dt \; \psi_j^\outp \enskip,
%\end{equation}
%where the interface states, $p_0^{n+\myhalf}$, are again found via the procedure 
%described in paper II, Appendix A.\\
%
%\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
%\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp}, 
%\psi^{n+\myhalf,\outp}]$:
%
%First use a predictor-corrector formulation for the pressure update in order to improve 
%the overall accuracy.  Beginning with equation (\ref{eq:w01}), we can write
%\begin{equation}
%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
% \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) 
%- \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
%\end{equation}
%We time-center the left hand side:
%\begin{equation}
%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
%\frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} 
%- p_0^{\inp}}{\dt}\enskip.
%\end{equation}
%We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
%\begin{equation}
%p_0^{\star} = p_0^{\inp} 
%\frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
%\end{equation}
%with:
%\begin{equation}
%f = \frac{\dt}{2} \left [ \Sbar^{\inp} - 
%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
%\frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} 
%\left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
%\end{equation}
%where the $(\partial p_0^\inp / \partial r)$ is computed by interpolating $p_0^\inp$ from 
%cell-centers to edges and then differencing.  Next, we compute a provisional updated 
%$\gammabar$ using the provisional updated pressure,
%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
%where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
%$X_k$ which were used to compute $\gammabar^\inp$.\\
%
%Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
%we can compute a more accurate approximation to equation (\ref{eq:frhs}),
%\begin{equation}
%f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -
%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
%\frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} 
%\left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
%\end{equation}
%where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$ and
%$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$.
%This allows us to complete the corrector for the pressure update:
%\begin{equation}
%p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\dagger} f^\dagger\right)}{\left(1 - \gammabar^{\dagger} f^\dagger\right)}\enskip .
%\end{equation}
%Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  First, we compute an 
%updated $\gammabar$ using the update pressure, 
%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^{\outp},X_k^{\inp})]$.  Then, 
%rewriting equation (\ref{eq:w01}), we can define
%\begin{eqnarray}
%\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
%       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
%\end{eqnarray}
%We discretize this locally as:
%\begin{equation}
%\psi_j^{\outp,n} = 
%\left(\gammabar^\inp p_0^\inp\right)_j
%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
%\end{equation}
%\begin{equation}
%\psi_j^{\outp,n+\myhalf} = 
%\left(\frac{\gammabar^\inp+\gammabar^{\star}}{2}\right)_j
%\left(\frac{p_0^\inp+p_0^\outp}{2}\right)_j
%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
%\end{equation}
%\end{description}
\item {\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp},g^{\inp}] \rightarrow [p_0^{\outp}]$
has replaced {\bf Advect Base Pressure} from paper III as the process by which we 
update the base state pressure.  Rather than discretize the evolution equation
for $p_0$, we enforce hydrostatic equilibrium directly, which is simpler and
analytically equivalent.  We update $p_0$ using:
\begin{equation}
p_{0,j}^{\outp} = p_{0,j-1}^{\outp} + 
\frac{\Delta r}{2}\left(\rho_{0,j}^{\inp}+\rho_{0,j-1}^{\inp}\right)g_{j-\myhalf}^{\inp}\enskip .
\end{equation}
Once the density is less than the cutoff density, we set $p_{0,j}^\outp = p_{0,j-1}^\outp$.
We use the boundary condition, $p_{0,jmax}^\outp = p_{0,jmax}^\inp$.

\item {\bf Advect Base Enthalpy}
is the process by which we update the base state enthalpy through $\dt$ in time.
The input arguments are different for planar and spherical geometries.
\begin{description}
\item {\bf Advect Base Enthalpy Planar}$[(\rho h)_0^\inp,w_0^\inp,\psi^\inp] \rightarrow 
[(\rho h)_0^\outp]$:

We discretize the base state enthalpy equation,
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states $(\rho h)_0^{n+\myhalf}$, are again found via the procedure 
described in paper II, Appendix A.\\

\item {\bf Advect Base Enthalpy Spherical}$[(\rho h)_0^\inp,w_0^\inp,\psi^{\inp,n},
\psi^{\inp,n+\myhalf}] \rightarrow [(\rho h)_0^\outp]$:

>>>>

Here we outline the algorithm currently implemented in the code.  Our
starting point for this description are the series of papers describing
the development of the algorithm:
\begin{itemize}
\item {\em Low Mach Number Modeling of Type Ia
  Supernovae. I. Hydrodynamics,} A. S. Almgren, J. B. Bell, 
  C. A. Rendleman, \& M. Zingale 2006, ApJ, 637, 922 (henceforth
  paper I)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. II. Energy
  Evolution,} A. S. Almgren, J. B. Bell, C. A. Rendleman, \& M. Zingale
  2006, ApJ, 649, 927 (henceforth paper II)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. III. Reactions,}
  A. S. Almgren, J. B. Bell, A. Nonaka, \& M. Zingale
  2008, ApJ, 684, 449 (henceforth paper III)
\item {\em Low Mach Number Modeling of Type Ia Supernovae. IV. White Dwarf Convection,}
  M. Zingale, A. S. Almgren, J. B. Bell, A. Nonaka, \& S. E. Woosley
  2009, ApJ, 704, 196 (henceforth paper IV)
\item {\em MAESTRO: An Adaptive Low Mach Number Hydrodynamics Algorithm for Stellar
  Flows,} A. Nonaka, A. S. Almgren, J. B. Bell, M. J. Lijewski, C. M. Malone,
  \& M. Zingale 2009, submitted to SIAM J. Sci. Comput. 
  (henceforth ``the multilevel paper'')
\end{itemize}
We also carry over some ideas from our small-scale low Mach number algorithm
for astrophysical flows:
\begin{itemize}
\item {\em Adaptive Low Mach Number Simulations of Nuclear Flames,}
J. B. Bell, M. S. Day, C. A. Rendleman, S. E. Woosley, \& M. Zingale
2004, JCP, 195, 2, 677 (henceforth BDRWZ)
\end{itemize}

\section{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item We defined the mapping of data between a 1D radial array and the 3D Cartesian
grid for spherical problems (which we improve upon in the multilevel paper).
\item We update $T$ after the call to {\bf React State}.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We use {\tt use\_tfromp} to update temperature using $T=T(\rho,X_k,p_0)$ rather 
than $T=T(\rho,h,X_k)$.
\item For spherical problems, we have changed the discretization of 
$\Ubt\cdot\nabla p_0$ in the enthalpy update to 
$\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
\item In paper III we discretized the enthalpy evolution equation in
terms of $T$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\end{enumerate}
\section{Changes Between Paper 4 and the Multilevel Paper}
See the multilevel paper for the latest.
\section{Changes Between Papers 4.5 and Paper 5}
\begin{enumerate}
\item Added rotation.
\end{enumerate}
\section{Changes Between Paper 5 and XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.
\end{enumerate}

@@

$c_p$   & specific heat at constant pressure
          ($c_p \equiv \left . \partial h / \partial T \right |_{p,X_k}$)
        & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$f$     & volume discrepancy factor ($0 \le f \le 1$) & -- \\
\hline
$g$     & gravitational acceleration                 & cm~s$^{-2}$ \\
\hline
$h$     & specific enthalpy                          & erg~g$^{-1}$ \\
\hline
$\Hext$ & external heating energy generation rate    & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$\Hnuc$ & nuclear energy generation rate             & erg~g$^{-1}$~s$^{-1}$ \\
\hline
$h_p$   & $h_p \equiv \left . \partial h / \partial p \right |_{T,X_k}$ & cm$^{3}$~g$^{-1}$ \\
\hline
$\kth$  & thermal conductivity                       & erg~cm$^{-1}$~s$^{-1}$~K$^{-1}$ \\
\hline
$p_0$   & base state pressure                        & erg~cm$^{-3}$ \\
\hline
$p_T$   & $p_T \equiv \left . \partial p / \partial T \right |_{\rho,X_k}$ & erg~cm$^{-3}$~K$^{-1}$ \\
\hline
$p_{X_k}$ & $p_{X_k} \equiv \left . \partial p / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~cm$^{-3}$ \\
\hline
$p_\rho$ & $p_\rho \equiv \left . \partial p / \partial \rho \right |_{T,X_k}$ & erg~g$^{-1}$ \\
\hline
$q_k$   & specific nuclear binding energy            & erg~g$^{-1}$  \\
\hline
$r$     & radial coordinate (direction of gravity)   & cm \\
\hline
$s$     & specific entropy                           & erg~g$^{-1}$~K$^{-1}$ \\
\hline
$S$     & source term to the divergence constraint   & s$^{-1}$ \\
\hline
$t$     & time                                       & s \\
\hline
$T$     & temperature                                & K \\
\hline
$\Ub$     & total velocity ($\Ub = \Ubt + w_0 \eb_r$) & cm~s$^{-1}$ \\
\hline
$\Ubt$   & local velocity                             & cm~s$^{-1}$ \\
\hline
$\uadv$ & advective velocity (edge-centered)         & cm~s$^{-1}$ \\
\hline
$w_0$   & base state expansion velocity              & cm~s$^{-1}$ \\
\hline
$X_k$   & mass fraction of the species ($\sum_k X_k = 1$) & -- \\
\hline
$\beta_0$ & coefficient to velocity
            in velocity constraint equation  & g~cm$^{-3}$ \\
\hline
$\Gamma_1$ & first adiabatic exponent ($\Gamma_1 \equiv \left . d \log p/d \log \rho \right |_s$) & -- \\
\hline
$\etarho$ & $\etarho \equiv \overline{(\rho' \Ub \cdot \eb_r)}$ & g~cm$^{-2}$~s$^{-1}$ \\
\hline
$\xi_k$ & $\xi_k \equiv \left . \partial h / \partial X_k \right |_{p,T,X_{j,j\ne k}}$ & erg~g$^{-1}$ \\
\hline 
$\pi$   & dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\pizero$ & base state dynamic pressure & erg~cm$^{-3}$ \\
\hline
$\rho$  & mass density  & g~cm$^{-3}$ \\
\hline
$\rho_0$  & base state mass density  & g~cm$^{-3}$ \\
\hline
$\rho'$  & perturbational density ($\rho' = \rho - \rho_0$) & g~cm$^{-3}$ \\
\hline
$(\rho h)_0$ & base state enthalpy density & erg~cm$^{-3}$  \\
\hline
$(\rho h)'$ & perturbational enthalpy density 
              $ \left [(\rho h)' = \rho h - (\rho h)_0 \right ]$ & erg~cm$^{-3}$  \\
\hline
$\sigma$ & $\sigma \equiv p_T/(\rho c_p p_\rho)$ & erg$^{-1}$~g \\
\hline
$\psi$  & $\psi \equiv D_0 p_0/Dt = \partial p_0/\partial t + w_0\partial p_0/\partial r$ & erg~cm$^{-3}$~s$^{-1}$ \\
\hline
$\omegadot_k$ & creation rate for species $k$ ($\omegadot_k \equiv DX_k/Dt$) & s$^{-1}$ \\
\end{longtable}
\end{center}

\renewcommand{\arraystretch}{1.0}
There are some great charts in the appendix of the multilevel paper demonstrating what 
the superscript notation represents.
\section{Shorthand Functions}
See the multilevel paper for the latest.
